MSE and dynamic state variable model with sigma function

### 2017-10-03 doc/MSE.tex
changes in text and scenarios

\documentclass[12pt,oneline,a4paper,numbib]{ouparticle}

@@

The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation to encourage more selective fishing practices. Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other.

@@

Using a MSE, the main objective of this work is to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.

@@

Mortality in the model can occur from catches in the fishery, and natural causes such as predation, diseases, and scenesens. However, for simplicity we assume that his natural mortality is negligible. Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species. The decrease in population numbers is thus the result of the catches, which are in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.

@@

Migration for each species is defined by a matrix that defines the migration to and from each of the areas in the model. Its dimensions are thus equal to the number of areas in the model.

the natural mortality at age $a$ in seson $s$ and $F (a, y, s, q)$ is the fishing mortality at age $a$ in year $y$, season $s$ and area $q$. $p (a,y,s,i, q)$ is the proportion of the population available or vulnerable to fishing after migration occurs. Life history traits inputted are: (i) each season less than 1\% of specie 1 and 2 die of natural causes, (ii)  during the first season, specie 1 and 2 spawn in ``A'', (iii) every season, 20\% of specie 1 and 2 migrate from area ``A'' to ``B'' and from ``B'' to ``A'', (iv) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (v) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (vi) at the beginning of each year the fishing recruitment is stablish in 100.

@@

We develop models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming (Clark and Mangel, 2000; Houston and McNamara, 1999; Poos et al., 2010; Dowling et al., 2011; Batsleer et al., 2015 and Alzorriz et al., 2017).

In order to calculate the optimal state dependent choices during the year, we start by defining the contribution of the annual fines for exceeding landings quotas at the end of the year to the utility is: NEED TO CHECK LANDINGS or CATCH!!
\begin{equation*}
\Phi (L_i, Q_i^L, F_i^L) = -\sum_{i} (max(0, (L_i - Q_i^L)\times F_i^L),
\end{equation*}

where $L_i$ are the cumulative annual landings, for species i. $Q_i^L$ are the annual quota for landings. $F_i^L$ are the fines per unit weight for exceeding landings quota. Hence, starting with $V (L, Q, F, t) = \Phi (L, Q, F)$ we assigned a probability to each region patch proportional to its profit associated at that point in the season (t), given the quota remaining. Following Dowling et al., 2011, if $V^\ast$ is the profit at the optimal location for a given Q and t, we set
\begin{equation*}
\Delta (Q_i^L, t) = V^\ast (L, Q, F, t) - V (L, Q, F, t),
\end{equation*}

and then define the probability of fishing particular area as
\begin{equation*}
P (Q_i^L, t) = \frac{e^-\Delta (Q_i^L, t)/\sigma} {\sum_{i}(e^-\Delta (Q_i^L, t)/\sigma)},
\end{equation*}

where $\sigma$ is a tuning parameter that measures how important is to be near optimal.  A high value of sigma would mean that probabilities are uniform and vessels are distributed uniformly along the different fishing areas, while a low sigma value would mean that vessels concentrate in the optimal location. For computations, we use $\sigma = 40$, to introduce some error in the decision making.

Then, we can iterate backwards in time and fine the optimal choice in terms of location and discarding behaviour for all possible states, combinig the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding anuual quota.

\begin{table}[!h]
\centering
\caption{Fishery parameterization.}
\label{t:fishery}
\resizebox{0.4\textwidth}{!}{%
\begin{tabular}{@{}lc@{}} 
\hline
\textbf{model}                              &   \\ \hline
\\
$\sigma $                                         & 40\\
number of areas & 2 \\
number of seasons                                           & 6\\
\\
\hline
\textbf{Variable costs}                              &\\ \hline
\\
fuel costs (Euro day\textsuperscript{-1})       & 1240\\
gear maintenance (Euro day\textsuperscript{-1}) & 287 \\
crew share                                           & 35\%\\
landing costs (Euro t\textsuperscript{-1})      & 121\\
\\
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}&  \\ \hline
                      \\
sp1                  & 1000 \\
sp2                  & 1000   \\

\\
\hline 
\end{tabular}
}
\end{table}

- Prices of fish have been assumed to be constant over fish age and time.

@@

@@

estimating the hr Fisheries are managed by a total allowable catch (TAC). A
maximum TAC is set for each stock so that the respective target biomass is maintained
on average. This maximum TAC may be taken as long as biomass fluctuations remain
above 
B
msy
.

@@

Simulations have been performed under four different scenarios. All compared to an unmanaged fishery,  with unconstrained quota for both species.

@@

\item
Scenario I,  decrease the quota on both species to reach the Fmsy.
Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).

\item 
Scenario II, decrease the quota on sp 1 and unconstrained quota for species 2.
CONSTRAINED fishery for sp1 and UNCONSTRAINED sp2 (FIXING QUOTA SP2 to 200). Only sp1 is forced to try to reach the Fmsy. However discarding is allowed for both species.

\item
Scenario III, same as scenario II but allowing to discard only species 2.

\item
Scenario IV, Same as scenario I but allowing to discard sp2.
\end{enumerate}

@@

@@

\bibliographystyle{plain}
\bibliography{mylib}

>>>>

\documentclass[12pt,oneline,a4paper,numbib]{ouparticle}
\usepackage{array,multirow,graphicx}
\usepackage{pdflscape}% for landscape

@@

The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation (LO) to encourage more selective fishing practices. Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other. LO and catch quotas should provide incentives for change in the fisheries, including adoption through taking up selective gears and spatiotemporal effort allocation.

@@

Using a MSE, the main objective of this work is to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

@@

Mortality in the model can occur from catches in the fishery, and natural causes such as predation, diseases, and scenesens. However, for simplicity we assume that his natural mortality is negligible. Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species. The decrease in population numbers is thus the result of the catches ($C (a, y, s, p)$), which are in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.

@@

Migration for each species is defined by a matrix that defines the migration to ($top$) and from ($fromp$) each of the areas in the model. Its dimensions are thus equal to the number of areas in the model.  The migration in population numbers ($Mig (a, 1, s, top, fromp)$) is thus the proportion ($p (a, y, s, i, fromq, top)$) of the population available or vulnerable to fishing after migration occurs.   
\begin{equation*}
Mig (a, 1, s, top, fromp) = Mig(a, 1, s, top, fromp) + N (a, y, s, fromp) * p (a, y, s, i, fromq, top)
\end{equation*}

and then population numbers
\begin{equation*}
N (a, y, s, top) = N (a, y, s, top) +  Mig (a, 1, s, top, fromp)
\end{equation*}

Life history traits inputted are: (i)  during the first season, specie 1 spawns in ``A'', while specie 2 spawns in ``B'', (ii) every season, 20\% of specie 1 and 2 migrate from area ``A'' to ``B'' and from ``B'' to ``A'', (iii) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (iv) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (v) at the beginning of each year the fishing recruitment is stablish in 100.

@@

We develop models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2017, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.

In order to calculate the optimal state dependent choices during the year, we start by defining the contribution of the annual fines for exceeding landings or catches quotas at the end of the year to the utility is: 
\begin{equation*}
\Phi (C_i, Q_i^C, F_i^C) = -\sum_{i} (max(0, (C_i - Q_i^C)\times F_i^C),
\end{equation*}

where $C_i$ are the cumulative annual catches, for species i. $Q_i^C$ are the annual quota for catches. $F_i^C$ are the fines per unit weight for exceeding catch quota. Hence, starting with $V (C, Q, F, t) = \Phi (C, Q, F)$ we assigned a probability to each region patch proportional to its profit associated at that point in the season (t), given the quota remaining. Following \cite{Dowling2011}, if $V^\ast$ is the profit at the optimal location for a given Q and t, we set
\begin{equation*}
\Delta (Q_i^C, t) = V^\ast (C_i, Q_i^C, F_i^C, t) - V (C_i, Q_i^C, F_i^C, t),
\end{equation*}

and then define the probability of fishing particular area as
\begin{equation*}
P (Q_i^C, t) = \frac{e^-\Delta (Q_i^C, t)/\sigma} {\sum_{i}(e^-\Delta (Q_i^C, t)/\sigma)},
\end{equation*}

where $\sigma$ is a tuning parameter that measures how important is to be near optimal.  A high value of sigma would mean that probabilities are uniform and vessels are distributed uniformly along the different fishing areas, while a low sigma value would mean that vessels concentrate in the optimal location. For computations, we use $\sigma = 40$, to introduce some error in the decision making (Table \ref{t:fishery}).

Then, we can iterate backwards in time and fine the optimal choice in terms of location and discarding behaviour for all possible states, combinig the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding anuual quota. Prices of fish have been assumed to be constant over fish age and time. The variable costs include fuel costs ($\sim{40\%}$), crew share ($\sim{35\%}$) and gear mantainance ($\sim{4\%}$)). Fuel costs depend on gear used, trip effort and fuel price. Fuel costs were set to 150 \texteuro day\textsuperscript{-1}. Gear maintenance costs were assumed proportional to fishing effort, landing costs proportional to the total weight landed, and crew share variable costs proportional to the income from catches (Table~\ref{t:fishery}).

\begin{table}[!h]
\centering
\caption{Fishery parameterization.}
\label{t:fishery}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}|lc|lc|lc|@{}} 
\hline
\textbf{model}                              &               & \textbf{Variable costs}                        &      &\textbf{Market value}&  \\ \hline
& & & & & \\
number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  & sp1 (Euro kg\textsuperscript{-1})                 & 1000 \\ 
$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 287  & sp2 (Euro kg\textsuperscript{-1})                 & 1000   \\
number of areas                             & 2             &crew share                                      & 35\% & & \\
number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    & &\\
intial quota $Q_sp1$ and $Q_sp2$            & 200           &                                                &      & & \\
& & & & & \\\hline
\end{tabular}
}
%\hline
%\textbf{model}                              &   \\ \hline
%\\
%number of vessels                                 &700\\
%$\sigma $                                         & 40\\
%number of areas                                   & 2 \\
%number of seasons                                 & 6\\
%intial quota Q1 and Q2                            & 200\\
%\\\hline
%\textbf{Variable costs}                              &\\ \hline
%\\
%fuel costs (Euro day\textsuperscript{-1})       & 1240\\
%gear maintenance (Euro day\textsuperscript{-1}) & 287 \\
%crew share                                      & 35\%\\
%landing costs (Euro t\textsuperscript{-1})      & 121\\
%\\\hline
%\textbf{Market value (Euro kg\textsuperscript{-1})}&  \\ \hline
%\\
%sp1                  & 1000 \\
%sp2                  & 1000   \\
%\\\hline 
%\end{tabular}
%}
\end{table}

% Prices of fish have been assumed to be constant over fish age and time.

@@

prel.quota <-  sum(sweep((hr1wanted[1]/mean(hr1[,yy,]))* hr1[,yy,]*landings.ratio1[,yy,]*apply(pop1[,yy+1,,],c(1,2), sum) ,1,wts,"*"))/SIMNUMBER
        tac.constrained <- c(quota1[,yy,,]*0.85, quota1[,yy,,]*1.15)
        quota1[,yy+1,,] <- max(min(prel.quota, tac.constrained[2]), tac.constrained[1])

@@

estimating the hr Fisheries are managed by a total allowable catch (TAC). A maximum TAC is set for each stock so that the respective target biomass is maintained
on average. This maximum TAC may be taken as long as biomass fluctuations remain above Bmsy.

@@

Simulations have been performed under three different scenarios. All compared to an unmanaged fishery, with unconstrained quota for both species.

\begin{table}[!h]
\centering
\caption{}
\label{scenarios}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|c |c | c | c |}
\hline
 \textbf{Scenario I}            & \textbf{Scenario II}                & \textbf{Scenario III}               & \textbf{ScenarioIV}           \\ 
 \textbf{Fulll avoidance of discards} & \textbf{Landings selectivity} & \textbf{Catch selectivity}          & \textbf{LO not fully enforced} \\ \hline
 & &  &\\
 No discarding allowed          & Discarding allowed                  & Discarding allowed                  & Discarding ocurred but not perceived \\
 Manager perceives $C$          & Manager perceives $L$ and $D$       & Manager perceives $L$ and $D$       & Manager only perceives $L$ \\
 assuming that $C = L$          & assuming that $C = L + D$           & assuming that $C = L + D$           & assuming that $C = L$  \\
 YPR based in $F_C$             & YPR based in $F_L$                  & YPR based in $F_C$                  &  YPR based in $F_L$  \\
 assuming $L/C=1$               & assuming $L/C \neq 1$               & assuming $L/C \neq 1$               & assuming $L/C=1$\\
  & &  &\\\hline
\end{tabular}
}
\end{table}

% \parbox[c]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{\textbf{Naive manager}}}} 
% & &  \\
 %& Discarding allowed & \\
% & Manager only perceives $L$ &\\
% & assuming that $C = L$  &\\
% & YPR based in $F_L$  & \\
% & assuming $L/C=1$    & No discarding allowed\\
% & & $C = L$ \\\cline{1-2}
%  \parbox[c]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{\textbf{Smart manager}}}} 
% & &  Advice and quota\\
% & Discarding allowed & based on $C = L$\\
% & Manager perceives $L$ and $D$ & \\
% & assuming that $C = L + D$ &\\
% & YPR based in $F_C$  &\\
% & assuming $L/C \neq 1$ &\\
%  & &  \\\hline
% & \textbf{NO LO}              & \textbf     {LO}                                  \\ \hline

@@

\item
\textbf{Scenario I: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
%decrease the quota on both species to reach the Fmsy.
%Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).

\item
\textbf{Scenario II: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield. Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.

\item
\textbf{Scenario III: Catch selectivity},  under this scenario the fishery is under a full catch selectivity, but all discards are landed and contribute to yield. Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded. This results generally in higher $F_{MSY}$.
%decrease the quota on both species to reach the Fmsy.
%Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).

\item 
\textbf{Scenario IV: LO not fully enforced}, if LO is not fully enforced, therefore, there will be the incentive to continue discarding. However the manager could think that is fully enforced and assumed that only what is landed contribute to yield. This is the typicall situation of a choke species, where the management target such as $F_{MSY}$ cannot be easily achieved, because reductions in TACS do not necessarily translate into the expected reduction in fishing mortality. Rather, they may translate into increased discards if the fishery is mixed and continues fishing for other species beyond the exhaustion of the TAC of the species in question.
en vez de plantearlo como que no es LO, podria plantearse como que realmente esta bajo el LO, pero no hay compliance y el manager no lo sabe. Por tanto los hr se estiman como si el C fueran L...
%decrease the quota on sp 1 and unconstrained quota for species 2.
%CONSTRAINED fishery for sp1 and UNCONSTRAINED sp2 (FIXING QUOTA SP2 to 200). Only sp1 is forced to try to reach the Fmsy. However discarding is allowed for both species.

%same as scenario II but allowing to discard only species 2.

%\item
%Scenario IV, Same as scenario I but allowing to discard sp2.
\end{enumerate}

@@

\newpage
\begin{landscape}
Scenario I:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO.eps} 
\caption{}
\label{fig:catch_lo}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO.eps} 
\caption{}
\label{fig:effort_lo}
\end{figure}
\end{landscape}

%----------------------------------------------------------------------------------------------------------------------------------------
\newpage
\begin{landscape}
Scenario II:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO_disc_YieldLand.eps} 
\caption{}
\label{fig:catch_lo}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO_disc_YieldLand.eps} 
\caption{}
\label{fig:effort_lo}
\end{figure}
\end{landscape}

%----------------------------------------------------------------------------------------------------------------------------------------
\newpage
\begin{landscape}
Scenario III:
%----------------------------------------------------------------------------------------------------------------------------------------

\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO_smartmanager.eps} 
\caption{}
\label{fig:catch_lo}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO_smartmanager.eps} 
\caption{}
\label{fig:effort_lo}
\end{figure}
\end{landscape}

%----------------------------------------------------------------------------------------------------------------------------------------
\newpage
\begin{landscape}
Scenario IV:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO_nocompliance.eps}  
\caption{}
\label{fig:catch_lo}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO_nocompliance.eps} 
\caption{}
\label{fig:effort_lo}
\end{figure}
\end{landscape}
%----------------------------------------------------------------------------------------------------------------------------------------

@@

\newpage
\bibliographystyle{plain}
\bibliography{mylib}
